---
title: "Why is the Amazon burning?"
author: "Mirza Cengic"
date: "August 24, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

Development of satellite technology and earth observation methods has enabled us an unprecedented insight into the state of the planet Earth. There are currently over XXX earth observation satellites orbiting the earth, and we can access data from many of them for free. They give us information about ..., ..., ..., and they enable us to see the state of Earth's systems in very short amount of time, especially when time is crucial. One of the events that many seem to write about recently are the fires in South America, and in particular the fires in the Amazon rainforest.
Researchers have detected an increase in deforestation in Brazil, which increased since June. This has been a result of new policies and new sentiment of current brazil government.

There are many different reports coming from various sources. And the current state of information access in the current political climate of Brazil seems to be heading towards the road that Trump paved, where he restrained powers of the EPA. Since there are a lot of very useful data in the public domain, I decided to look for myself what is happening in South America.


The natural areas of South America, and the indigenous lands are under threat. Current president of Brazil, Jair Bolsonaro, considers indigenous people a "problem".
“The Brazilian cavalry was very incompetent. Competent, yes, was the American cavalry that decimated its Indians in the past and nowadays does not have this problem in their country.”

On his first day as president, Bolsonaro transferred the authority to protect Indigenous lands from Brazil’s National Indian Foundation, or FUNAI, a government entity tasked with the protection of Indigenous communities, to the Ministry of Agriculture, handing a victory to the powerful agribusiness sector that backed his campaign and has its eyes on large tracts of pristine forest. Sydney Possuelo, a veteran Indigenous observer and former FUNAI president, described the move as “the death” of FUNAI, in an interview with the Folha de São Paulo newspaper.


Note: FUNAI is an organization that sucedded SPI (Servico de Protecao ao Indio), established by Candido Rondon.

Bolsonaro won in Rondonia by a wide margin. Rondonia was named after Candido Rondon, who fought for the rights fo indigenous people.


https://theintercept.com/2019/02/16/brazil-bolsonaro-indigenous-land/
[link](https://piaui.folha.uol.com.br/lupa/2018/12/06/verificamos-bolsonaro-cavalaria/)


Find indigenous people territory shapefile. There could be a link here - [link-in](http://www.landmarkmap.org/data/#data-4)

https://unearthed.greenpeace.org/2017/12/01/brazil-amazon-indigenous-violence-rondonia/

Check how close are the fires to the 163 road leading to Miritituba.

[excellent article](https://theintercept.com/2019/07/06/brazil-amazon-rainforest-indigenous-conservation-agribusiness-ranching/)
[another one](https://theintercept.com/2019/08/23/gop-lobbyists-help-brazil-recruit-u-s-companies-to-exploit-the-amazon/)
https://phys.org/news/2019-08-amazon_1.html

[vizzuality article](https://medium.com/vizzuality-blog/the-amazon-is-on-fire-is-it-worse-than-normal-5fa430a7880e)

[Lovejoy article](https://advances.sciencemag.org/content/4/2/eaat2340)

[budget cuts](https://www.axios.com/amazon-rainforest-bolsonaro-brazil-south-america-fda08c12-0de2-43ce-93f0-d02709b18e7b.html)

[rudy guilani](https://theintercept.com/2018/10/05/rudy-giuliani-amazon-contract-brazil-election/)

[august and tipping points](https://www.newscientist.com/article/2214734-its-officially-now-the-worst-ever-august-for-amazon-deforestation/?utm_term=Autofeed&utm_medium=SOC&utm_source=Twitter#Echobox=1567101051)

[fires worldwide](https://www.nytimes.com/2019/08/28/climate/fire-amazon-africa-siberia-worldwide.html?smid=tw-nytimes&smtyp=cur#commentsContainer)


#### Script setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(Rahat, raster, tidyverse, extdplyr, lubridate, ggmap,  ggthemes, sf,  mapview, janitor, tictoc, fasterize, patchwork, glue, cowplot)

```



#### Load data 

```{r load_data}
# Loading data:
# VIIRS data, at 375meters, last 7 days and historical (from when?)
# MODIS data, at 1 km, last 7 days and historical
# ESA land cover data
# Administrative borders of Brazil
# Intact forests raster layer (Potapov)
# Amazon biome limits (polygon)
# Distance layers (not entirely done yet)
files_loaded_name <- "tmp/amazon_data_loaded.RData"



if (!file.exists(files_loaded_name))
{
# historical_modis <- st_read("Data/Modis_historical.gpkg")
historical_viirs <- st_read("Data/VIIRS_historical.gpkg")

# current_modis <- st_read("Data/Modis_current.gpkg")
current_viirs <- st_read("Data/VIIRS_current.gpkg")

# Land coverf
esa_brazil <- raster("Data_raw/ESA_Brazil.tif")

# Load shapfile with brazil states
brazil_adm <- st_read("Data/Brazil_GADM.gpkg") %>% 
  mutate(
    area = units::set_units(st_area(.), "km^2")
  )

# Intact forests layer
ifl_raster <- raster("Data/IFL_raster.tif")

# Load natural earth with NAs as non forest, cropped with amazon biome outline
NE_amazon_forest <- raster("Data/NE_samer_amazon.tif")


# Amazon biome limits
amazon_biome <- st_read("Data_raw/amapoly_ivb/amapoly_ivb.shp")
###################
# output_folder <- milkunize2("Projects/Other/Amazon")

# crops_mosaic_distance <- glue("{output_folder}/Crops_mosaic_distance_px.tif")

NE_samer_small <- raster("Data/NE_samer_small.tif")
# NE_samer_large <- raster("Data/NE_samer_large.tif")

samer_countries <- st_read("Data/samer_countries.gpkg") %>% 
    st_simplify()

# esa_1km <- "Projects/Other/Amazon/landcover_amazon_1km.tif" %>% 
  # milkunize2() %>% 
  # raster()
esa_1km <- raster("Data_raw/ESA_Brazil.tif")

# Indigenous areas
indigenous_areas <- st_read("Data/Indigenous_territories.gpkg")

save.image(file = files_loaded_name)
  
} else {
  load(files_loaded_name)
}

```


#### Density of fires per state

```{r}

# Do spatial join of VIIRS fires data, with the Brazil administrative borders data
# This is so we can see which fires are in Brazil, and in which state of Brazil (and how many)
current_viirs_brazil <- st_join(current_viirs, brazil_adm, join = st_intersects)

# Calculate number of fire events per state, and calculate the density of fire events per state
current_viirs_brazil_cleaned <- current_viirs_brazil %>% 
  filter(!is.na(NAME_0)) %>% 
  group_by(NAME_1) %>%
  mutate(events_n = n()) %>%
  ungroup() %>%
  transmute(
    date = ACQ_DATE,
    state = NAME_1,
    events_n,
    area,
    event_density = events_n / area
  ) 

# Calculate fire density per area
current_viirs_brazil_cleaned %>% 
  st_set_geometry(NULL) %>% 
  distinct(state, .keep_all = TRUE) %>% 
  arrange(desc(event_density)) %>% 
  dplyr::select(state, event_density)
```


https://www.nytimes.com/interactive/2019/08/24/world/americas/amazon-rain-forest-fire-maps.html



#### Calculate number of fires on 1 or 10km level
```{r}

fire_freq_filename <- "Data/Fire_freq.gpkg"
my_mask <- aggregate(NE_samer_small, 0.3 / res(NE_samer_small)[1])

tic("Raster to poly")
mask_1deg <- rasterToPolygons(my_mask) %>% 
  st_as_sf() %>% 
  transmute(
    vals = 1:n()
  )
toc()

tic("Spatial join")

# fire_frequency <- st_join(mask_1deg, current_viirs_brazil, join = st_intersects)
fire_frequency <- st_join(mask_1deg, current_viirs, join = st_intersects)

toc()


if (!file.exists(fire_freq_filename))
{
  



# fires_count <- 
tic("Count")
fire_freq_count <- fire_frequency %>%
  group_by(vals) %>% 
    summarize(n = n()) %>% 
  st_centroid() %>% 
  filter(n != 1)
toc()

# fire_freq_count %>% 
#   ggplot() + 
#   geom_sf(aes(size = n), color = "red")

st_write(fire_freq_count, fire_freq_filename)
  
} else {

  fire_freq_count <- st_read(fire_freq_filename)  
}

```

#### Fire frequency with forests burnt percentage
```{r}
# Needs work, some old paths and some useful extract functions

# lc <- "Data_raw/ESA_Brazil_1km.tif" %>% 
  # raster()


current_viirs_lc_filename <- "Data/current_viirs_lc_filename.gpkg"

if (!file.exists(current_viirs_lc_filename))
{
  
esa <- "ESA_landcover/TIFF/ESACCI-LC-L4-LCCS-Map-300m-P1Y-2015-v2.0.7.tif" %>% 
  milkunize2("data") %>% 
  raster()

tic("Extracting")
viirs_lc <- raster::extract(esa, current_viirs, na.rm = TRUE, sp = TRUE) %>% 
  st_as_sf() 
toc()

viirs_lc <- viirs_lc %>% 
  rename(lc = ncol(.) - 1) 
 
st_write(viirs_lc, current_viirs_lc_filename)

} else {
  
  viirs_lc <- st_read(current_viirs_lc_filename)
}

fire_frequency_lc <- st_join(viirs_lc, mask_1deg, join = st_intersects)
# fire_frequency_lc <- st_join(mask_1deg, viirs_lc, join = st_intersects)

# toc()

# fire_freq_count <- 
# aa <-   fire_frequency_lc %>% 
#     filter(!is.na(vals)) %>% 
#     mutate(lc = as.factor(lc))

forest_fire_pct <- fire_frequency_lc %>% 
  filter(!is.na(vals)) %>%
  pct_routine(vals, lc) %>% 
  filter(lc == 50) %>% 
  mutate(pct = pct * 100) %>% 
  ungroup()

grid_fire <- fire_frequency %>%
  group_by(vals) %>% 
    summarize(n = n()) %>% 
  # st_centroid() %>% 
  filter(n != 1)

points_fire_proportion <- grid_fire %>% 
  left_join(forest_fire_pct) %>% 
  mutate(
    pct = ifelse(is.na(pct), 0, pct)
  ) %>% 
  st_centroid() %>% 
   mutate(
      pct_rcl = as.factor(case_when(
        pct <= 1 ~ 0,
        pct > 1 & pct <= 25 ~ 25,
        pct > 25 & pct <= 50 ~ 50,
        pct > 50 & pct <= 75 ~ 75,
        pct > 75 & pct <= 100 ~ 100,
        TRUE ~ 0
      ))
    )


ggplot() +
  geom_sf(data = points_fire_proportion, aes(size = n, color = pct_rcl),
          # color = "red2",
            alpha = 0.7, inherit.aes = FALSE) +
  # scale_color_gradient(low = "#d73027", high = "#1a9850") +
  # scale_color_gradient2(low = "#f7fcb9", mid = "#addd8e", high = "#31a354") +
  geom_sf(data = amazon_biome, fill = "green", alpha = 0.1, color = "grey10") +
                      # geom_sf(data = samer_countries, fill = NA,
    # # linetype = "dotted",
    # color = "grey45") +
  coord_sf(datum = NA, xlim = c(-78, -42), ylim = c(-21, 12)) +
    theme_map() +
  theme(
    legend.position = "none"
  )





```


#### Fires map 

I can map another variable to dot color perhaps. Maybe percentage of forests as LC?

[map inspiration](https://www.hcn.org/articles/scientific-research-people-are-to-blame-for-longer-fire-seasons)
```{r}


# register_goowgle(key = "AIzaSyDwXacVpD7bbbdYKelR_rJzajdTaZw0gqI")


# bra <- c(left = -75, bottom = -30, right = -30, top = 5)

# myMap <- get_map(location = bra, maptype = "satellite",
                 # source = "google")

viirs_june <- historical_viirs %>% 
  filter(acq_date > as_date("2019-06-01")) %>% 
  rename(geometry = geom) 

fire_sample <- current_viirs_brazil %>% 
  sample_n(100)

ifl_resampled_filename <- "Data/IFL_raster_resampled.tif"

if (!file.exists(ifl_resampled_filename))
{
  ifl_raster_resampled <- resample(ifl_raster, NE_samer_small, method = "ngb")
  writeRaster(ifl_raster_resampled, ifl_resampled_filename, options = "COMPRESS=LZW")
} else {
  ifl_raster_resampled <- raster(ifl_resampled_filename)
}



# NE_samer_small <- raster("Data/NE_samer_small.tif")

r_stack <- stack(NE_samer_small, ifl_raster_resampled)

r_stack_agg <- aggregate(r_stack, 0.2 / res(NE_samer_small)[1])

NE_samer_small_gg <- raster_to_gg(r_stack_agg[[1]])
# NE_samer_large_gg <- raster_to_gg(NE_samer_large)
NE_amazon_forest_gg <- raster_to_gg(r_stack_agg[[2]])


# ifl_raster_gg <- raster_to_gg(ifl_raster)

p <- ggplot() +
      # geom_sf(data = current_viirs_brazil_cleaned, color = "red", size = 0.95,
  # geom_sf(data = ama) +
  geom_tile(data = NE_samer_small_gg, aes(x = x, y = y, fill = NE_samer_small),
            show.legend = FALSE) +
  scale_fill_gradient2(low = "grey60", 
                       mid = "lightyellow2",
                       high = "grey40") + # Grey
  # scale_fill_discrete() + # Grey
  geom_tile(data = NE_amazon_forest_gg, aes(x = x, y = y, fill = NE_samer_amazon), alpha = 0.45, fill = "green2") +
    # scale_colour_discrete(drop=FALSE) + #forces all levels to display
  # geom_sf(data = current_viirs_brazil_cleaned, color = "red2", size = 0.95,
  # geom_tile(data = ifl_raster_gg, aes(x = x, y = y, fill = IFL_raster), color = "green") +
  # geom_sf(data = fire_freq_count, aes(size = n), color = "red2",
  geom_sf(data = fire_sample, color = "red2", size = 0.95,
            alpha = 0.7, inherit.aes = FALSE) +
  geom_sf(data = amazon_biome, fill = "green", alpha = 0.1, color = "grey10") +
   annotate("text", x = -45.5, y = 2.2,
            label = "Point size indicates\nthe number of fires.",
            family = "ubuntu", size = 3.5, color = "grey10") +
    annotate("segment", x = -45, y = 1, 
             xend = -44.35, yend = -7.2, color = "grey10",
             arrow = arrow(angle = 20, length = unit(2, "mm"), type = "closed")) +
    # geom_sf(data = samer_countries, fill = NA,
    # # linetype = "dotted",
    # color = "grey45") +
  coord_sf(datum = NA, xlim = c(-78, -42), ylim = c(-21, 12)) +
    theme_map() +
  theme(
    legend.position = "none"
  )
p

# ggsave("Output/Figures/map_fireevents_amazon_hires.png", p, dpi = 300, width = 20, height = 16)


# ggsave("Output/Figures/map_firefreq_amazon_hires.png", p, dpi = 300, width = 20, height = 16)

p1 <- ggplot() +
      # geom_sf(data = current_viirs_brazil_cleaned, color = "red", size = 0.95,
  # geom_sf(data = ama) +
  geom_tile(data = NE_samer_small_gg, aes(x = x, y = y, fill = NE_samer_small),
            show.legend = FALSE) +
  scale_fill_gradient2(low = "grey60", 
                       mid = "lightyellow2",
                       high = "grey40") + # Grey
  geom_tile(data = NE_amazon_forest_gg, aes(x = x, y = y, fill = NE_samer_amazon), alpha = 0.45, fill = "green2") +
  # geom_sf(data = current_viirs_brazil_cleaned, color = "red2", size = 0.95,
  # geom_tile(data = ifl_raster_gg, aes(x = x, y = y, fill = IFL_raster), color = "green") +
  # geom_sf(data = points_fire_proportion, aes(size = n, color = pct), #THIS ONE
           # inherit.aes = FALSE) +
  scale_color_gradient(low = "#fd8d3c", high = "#bd0026") +
  # scale_color_gradient2(low = "#d73027", mid = "#ffffbf", high = "#1a9850") +
      geom_sf(data = samer_countries, fill = NA,
    # # linetype = "dotted",
    color = "grey40") +
  geom_sf(data = amazon_biome, fill = NA, alpha = 0.1, color = "grey10",
          size = 1) +
  coord_sf(datum = NA, xlim = c(-78, -42), ylim = c(-21, 12)) +
    theme_map() +
  theme(
    legend.position = "none"
  )

p1
# ggsave("Output/Figures/map_hires.png", p1, 
       # dpi = 400, width = 14, height = 10)

```


#### Data quality & landcover stuff

Confidence in fire events
```{r eval=FALSE, include=FALSE}
# 

fires_modis_lc <- raster::extract(esa_brazil, current_modis, na.rm = TRUE, sp = TRUE) %>% 
  st_as_sf() 


fires_viirs_lc <- raster::extract(esa_brazil, 
                                  current_viirs, 
                                  na.rm = TRUE, sp = TRUE) %>% 
  st_as_sf() 


fires_viirs_ifl_lc <- raster::extract(ifl_raster, 
                                  fires_viirs_lc, 
                                  na.rm = TRUE, sp = TRUE) %>% 
  st_as_sf() 



dist_mosaic_name <- glue("{output_folder}/mosaic_dist_wgs.tif")

mos_dist <- raster(dist_mosaic_name)
plot(mos_dist)

fires_viirs_ifl_dist_lc <- raster::extract(mos_dist, 
                                  fires_viirs_ifl_lc, 
                                  na.rm = TRUE, sp = TRUE) %>% 
  st_as_sf() 


(fires_viirs_ifl_dist_lc$mosaic_dist_wgs / 1000) %>% hist

fires_viirs_ifl_dist_lc %>% 
  mutate(
    distance = mosaic_dist_wgs / 1000
  ) %>% 
  filter(distance > 0) %>% 
  arrange(distance %>% desc()) %>% 
  slice(1:50) %>% 
  mapview()


fires_viirs_ifl_lc %>% 
  st_set_geometry(NULL) %>% 
  tabyl(ESA_Brazil, layer)

fires_viirs_lc
fires_modis_lc

mapview(NE_samer_small, maxpixels = ncell(NE_samer_small))

fires_viirs_ifl_lc %>% 
  rename(landcover = ESA_Brazil) %>% 
  tabyl(landcover) %>% 
  arrange(desc(n)) %>% 
  mutate(
    percent = percent * 100
  )

```





```{r wrangle, eval=FALSE, include=TRUE}
# rds <- roads_raw %>% 
  # filter(GP_RTP == 1)
# writeRaster(esa_brazil, "Projects/Amazon/Data/ESA_Brazil.tif", options = "COMPRESS=LZW")

res_1km <- (res(esa_brazil)[1] * 3)

res_1km / res(esa_brazil)[1]

esa_1deg_filename <- "Projects/Amazon/Data/ESA_Brazil_1deg.tif"


if (!file.exists(esa_1deg_filename))
{

  tic("1deg")
  lc_1deg <- raster::aggregate(esa_brazil, fact = 1 / res(esa_brazil)[1])
  toc()

  writeRaster(lc_1deg, esa_1deg_filename, options = "COMPRESS=LZW")
  
}

# tic("half degree")
# lc_05deg <- raster::aggregate(esa_brazil, fact = 0.5 / res(esa_brazil)[1])
# toc()

# tic("2 degree")
# lc_2deg <- raster::aggregate(esa_brazil, fact = 2 / res(esa_brazil)[1])
# toc()



esa_1km_filename <- "Projects/Amazon/Data/ESA_Brazil_1km.tif"

if (!file.exists(esa_1km_filename))
{

  tic("1km")
  lc_1km <- raster::aggregate(esa_brazil, fact = 3)
  toc()

  writeRaster(lc_1km, esa_1km_filename, options = "COMPRESS=LZW")
  
} else {
  lc_1km <- raster(esa_1km_filename)
}





mask_1deg <- rasterToPolygons(lc_1deg) %>% 
  st_as_sf() %>% 
  transmute(
    vals = 1:n()
  )

my_intersect <- st_intersection(x = mask_1deg, y = fires_viirs_lc)

fires_count <- my_intersect %>% 
  st_set_geometry(NULL) %>% 
  dplyr::select(vals, ESA_Brazil) %>% 
  group_by(vals) %>% 
  summarise(n = n()) %>% 
  ungroup() %>% 
  arrange(desc(n))

my_counts <- mask_1deg %>% 
  left_join(fires_count, by = "vals")


my_counts %>%
  # arrange(desc(n))
  filter(n < 250) %>% 
  mapview(zcol = "n")

plot(my_counts[2])
mapview(fires_lc, zcol = "ESA_Brazil")
mapview(my_counts, zcol = "n")# + mapview(fires_viirs_lc)

```

Category:
50 Treecover broadleaved, evergreen
120 Shrubland
180 Shrub or hebaceous cover
30 Mosaic cropland (>50% crops)
100 Mosaic tree and shrub (>50%)
130 Grassland
11 Herbaceous cover crops
61 Tree cover, broadleaved, closed


What is the most common type of land that is burning?
What are the frequencies?

Calculate distance from roads.
Calculate distance from current crops.
Calculate distance from urban areas.

#### Distance from features

```{r eval=FALSE, include=TRUE}
fires_viirs_lc

lc_1km
roads_raw
roads_rasterized <- "Projects/Amazon/Data/roads_rasterized.tif"

tic("Rasterizing roads")
  rasterize_call <- glue::glue("gdal_rasterize -tr {round(res_1km, 12)} {round(res_1km, 12)} -te -180 -57 180 84 -l GRIP_roads_subset -burn 1 -a_nodata NA -co compress=LZW {roads_filename} {roads_rasterized}")
  # system(rasterize_call)
  toc()

  

# roads_raster <- fasterize::fasterize(roads_raw, lc_1km)


```


#### Temporal trends


```{r}

viirs_date <- historical_viirs %>% 
  transmute(
    acq_date,
    period = str_c(month(acq_date), "_", year(acq_date))
  ) %>% 
  rename(geometry = geom)

viirs_date_freq <- viirs_date %>% 
  # sample_n(1000) %>% 
  group_by(period) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  distinct(period, .keep_all = TRUE)

viirs_date_freq %>% 
  # sample_n(10000) %>% 
  ggplot() +
  aes(x = acq_date, y = n) +
  # geom_line() +
  geom_col() +
  scale_x_date(date_labels = "%b %Y",
               date_breaks = "6 months",
               position = "bottom") +
  theme_minimal()
  NULL


```

#### Fire frequency per state

Break down the same over time per state
```{r}



historical_viirs_brazil_name <- "Data/historical_viirs_brazil.gpkg"

if (!file.exists(historical_viirs_brazil_name))
{
  historical_viirs_brazil <- st_join(viirs_date, brazil_adm, join = st_intersects)
  
  historical_viirs_brazil %>% 
    st_write(historical_viirs_brazil_name)
} else {
  historical_viirs_brazil <- st_read(historical_viirs_brazil_name)
}


historical_viirs_brazil_cleaned <- historical_viirs_brazil %>% 
  mutate(id = str_c(period, "_", NAME_1)) %>% 
  filter(!is.na(NAME_0)) %>% 
  group_by(NAME_1, period) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  distinct(id, .keep_all = TRUE)


historical_viirs_brazil_cleaned %>%
  distinct(period, .keep_all = TRUE) %>% 
  transmute(acq_date, NAME_1, n) %>% 
  ggplot() +
  aes(x = acq_date, y = n) +
  # geom_line() +
  geom_col(fill = "red2") +
  scale_x_date(date_labels = "%b %Y",
               date_breaks = "6 months",
               position = "bottom") +
  theme_minimal() +
  # facet_grid(~ NAME_1)
  NULL

my_dates <- historical_viirs_brazil %>% 
  st_set_geometry(NULL) %>% 
  distinct(acq_date)
  
my_states <- historical_viirs_brazil %>% 
  st_set_geometry(NULL) %>% 
  distinct(NAME_1) %>% 
  filter(!is.na(NAME_1)) %>% 
  pull() %>% 
  as.character()


historical_fires_df <- historical_viirs_brazil_cleaned %>% 
  st_set_geometry(NULL) %>% 
  transmute(acq_date, NAME_1, id, n)

fires_freq_brazil <- crossing(my_dates, my_states) %>% 
  mutate(
        period = str_c(month(acq_date), "_", year(acq_date)),
        id = str_c(period, "_", my_states)
  ) %>% 
  left_join(historical_fires_df, by = "id") %>% 
  transmute(
    id,
    acq_date = acq_date.x,
    state = my_states,
    period,
    fires_n = ifelse(is.na(n), 0, n)
  ) %>% 
  distinct(id, .keep_all = TRUE)


top_fire_states <- fires_freq_brazil %>% 
  group_by(state) %>% 
  summarize(n_sum = sum(fires_n)) %>% 
  ungroup() %>% 
  top_n(16, n_sum) %>% 
  pull(state)

fires_freq_brazil %>% 
  filter(state %in% top_fire_states) %>%
    ggplot() +
  aes(x = acq_date, y = fires_n) +
  # geom_line(color = "red") +
  geom_col(color = "red2") +
  scale_x_date(date_labels = "%Y",
               date_breaks = "1 year") +
  labs(x = NULL, y = NULL) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    strip.text = element_text(size = 16),
    axis.text.x = element_text(size = 10), 
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    # panel.grid.major.x = element_blank(),
    NULL
  ) +
  facet_wrap(~ state)
  

```

#### Indigenous territories

```{r}


indigenous_areas_brazil <- indigenous_areas %>%
  filter(country == "Brasil")


mydat <-  points_fire_proportion %>% 
  st_set_geometry(NULL)  %>% 
    sample_n(100)# %>%
    # pull(pct)

brks_scale <- levels(mydat$pct_rcl)
labels_scale <- str_c(rev(brks_scale), "%")
labels_scale2 <- c("76 - 100%", "51 - 75%", "26 - 50%", "1 - 25%", "0%")


colorscale <- c('#fc8d59','#ef6548','#d7301f','#b30000','#7f0000')



p_ind <- ggplot() +
  #### Basemap - natural earth
  geom_tile(data = NE_samer_small_gg, aes(x = x, y = y, fill = NE_samer_small),
            show.legend = FALSE) +
  #### Basemap colorscheme
  scale_fill_gradient2(low = "grey60",
                       mid = "lightyellow2",
                       high = "grey40") + # Grey
    #### IFL forest layer aggregated (faster plotting)
  geom_tile(data = NE_amazon_forest_gg, aes(x = x, y = y, fill = NE_samer_amazon), alpha = 0.45, fill = "green2") +
  #### Only brazil - events per state, and density
  # geom_sf(data = current_viirs_brazil_cleaned, color = "red2", size = 0.95,
  #### Full version of IFL (I guess)
  # geom_tile(data = ifl_raster_gg, aes(x = x, y = y, fill = IFL_raster), color = "green") +
 geom_sf(data = indigenous_areas_brazil, fill = "orange", alpha = 0.3, color = "grey1") +
  #### Fires proportional circles (color not mapped)
   # geom_sf(data = points_fire_proportion, aes(size = n, color = pct), 
        ## Smaller dataset (faster plotting)
         # geom_sf(data = fire_sample, color = "red2", size = 0.95,
         # alpha = 0.7, inherit.aes = FALSE) +
#### Amazon biome outline
    geom_sf(data = amazon_biome, fill = "green", alpha = 0.1) +
  #### South america country borders    
  # geom_sf(data = samer_countries, fill = NA,
    # color = "grey45") +
  #### Fires proprtional circles (colors mapped)
  geom_sf(data = points_fire_proportion, aes(size = n, color = pct_rcl),
          # color = "red2",
          inherit.aes = FALSE) +
  scale_color_manual(
          # in manual scales, one has to define colors, well, manually
          # I can directly access them using viridis' magma-function
          values = colorscale,
          breaks = rev(brks_scale)) +
      # scale_color_gradient(low = "#d7301f", high = "#7f0000") +
   coord_sf(datum = NA, xlim = c(-78, -42), ylim = c(-21, 12)) +
    theme_map() +
  theme(
    legend.position = "none"
  )

# p_ind

ggsave("Output/Figures/map_firefreq_indigenous.png", p_ind,
       dpi = 300, width = 16, height = 12)

ggsave("Output/Figures/map_firefreq_indigenous2.png", p_ind,
       dpi = 400, width = 16, height = 12)


colorscale_gg <- mydat %>% 
    # select(qq, pct) %>% head
    ggplot() +
    aes(x = 1, y = pct_rcl) +
    geom_tile(aes(fill = pct_rcl)) +
  theme(
    legend.margin=margin(c(1,5,5,5))
    ) +
   scale_fill_manual(
          # in manual scales, one has to define colors, well, manually
          # I can directly access them using viridis' magma-function
          values = colorscale,
          breaks = rev(brks_scale),
          labels = labels_scale2,
          drop = FALSE,
          guide = guide_legend(
            direction = "horizontal",
            keyheight = unit(5, units = "mm"),
            keywidth = unit(50 / length(labels), units = "mm"),
            title.position = "top",
            title = "Percentage of forests burnt",
            title.theme = element_text(size = 38, 
                                       # face = "italic",
                                       family = "AvantGarde"),
            # I shift the labels around, the should be placed 
            # exactly at the right end of each legend key
            title.hjust = 0.5,
            label.theme = element_text(size = 24, 
                                       # face = "italic",
                                       family = "AvantGarde"),
            # label.hjust = 1,
            nrow = 1,
            byrow = T,
            # also the guide needs to be reversed
            reverse = T,
            label.position = "bottom",
            NULL
            )
      ) +
  theme(
    plot.margin = margin(t = 10, r = 100, b = 10, l = 10, unit = "cm")
  )


library(ggpubr)
library(gridExtra)
library(grid)
colorscale_gg_legend <- ggpubr::get_legend(colorscale_gg)
plot.new()

blank <- rectGrob(gp=gpar(col="white"), width = 0.02)
grid.arrange(colorscale_gg_legend, pointscale_gg_legend, 
             ncol = 2
             )



ggsave("Output/Figures/legend_colorscale_map.png", colorscale_gg_legend,
       dpi = 300, width = 16, height = 6)


#### Create legend for proportional points

pointscale_gg <- points_fire_proportion %>% 
    mutate(
      n_rcl = as.factor(case_when(
        n <= 10 ~ "0 - 10",
        n > 10 & n <= 100 ~ "11 - 100",
        n > 100 & n <= max(n) ~ "101 - 500+",
        TRUE ~ as.character(n)
      )),
      n_num = case_when(
        n <= 10 ~ 10,
        n > 10 & n <= 100 ~ 100,
        n > 100 & n <= max(n) ~ 1000,
        TRUE ~ 0
      ),
      n_rcl = fct_relevel(n_rcl, levels = c("0 - 10", "11 - 100", "101 - 500+"))
    ) %>% 
    # tabyl(n_rcl)
  # select(qq, pct) %>% head
    ggplot() +
    aes(x = 1, y = n_rcl) +
  geom_point(aes(size = n_rcl), color = "#b30000") +
  guides(size = guide_legend(
    direction = "horizontal",
    keyheight = unit(5, units = "mm"),
    # keyheight = unit(10, units = "mm"),
    # keywidth = unit(10, units = "mm"),
    title = "Number of fires",          
    title.position = 'top',
    title.theme = element_text(size = 38, 
                               # face = "italic",
                               family = "AvantGarde"),
    # I shift the labels around, the should be placed 
    # exactly at the right end of each legend key
    # title.hjust = 0.5,
    # title.vjust = -1,
    label.theme = element_text(size = 24, 
                               # face = "italic",
                               family = "AvantGarde"),
    # label.hjust = 1,
    # nrow = 1,
    # byrow = T,
    # also the guide needs to be reversed
    # reverse = T,
            label.position = "bottom",
    NULL
    )
  
)

pointscale_gg_legend <- get_legend(pointscale_gg)
plot.new()
ggdraw(pointscale_gg_legend, xlim = c(0, 1))

ggsave("Output/Figures/legend_pointscale_map.png", pointscale_gg_legend,
       dpi = 300)#, width = 8, height = 6)


plot_grid(p_ind, colorscale_gg_legend, nrow = 2,
          rel_heights = c(2, 1)
          )



combined_leg <- (ggdraw(pointscale_gg_legend) | ggdraw(colorscale_gg_legend))

combined_leg

```

#### Plot short
```{r}

amazon_main <- "Data_raw/CD06_CAMREX_1086/data/amzmainstm_v/amzmainstm_v.shp" %>% 
  st_read() %>% 
  st_transform(crs = st_crs(amazon_biome))

# r_stack <- stack(NE_samer_small, ifl_raster_resampled)

NE_samer_small_ggg <- raster_to_gg(r_stack[[1]])
# NE_samer_large_gg <- raster_to_gg(NE_samer_large)
NE_amazon_forest_ggg <- raster_to_gg(r_stack[[2]])

plot_firedens <- ggplot() +
      # geom_sf(data = current_viirs_brazil_cleaned, color = "red", size = 0.95,
  # geom_sf(data = ama) +
  geom_tile(data = NE_samer_small_ggg, aes(x = x, y = y, fill = NE_samer_small),
            show.legend = FALSE) +
  scale_fill_gradient2(low = "grey60", 
                       mid = "lightyellow2",
                       high = "grey40") + # Grey
  geom_tile(data = NE_amazon_forest_ggg, aes(x = x, y = y, fill = ifl_raster_resampled), alpha = 0.45, fill = "green2") +
  # geom_sf(data = current_viirs_brazil_cleaned, color = "red2", size = 0.95,
  # geom_tile(data = ifl_raster_gg, aes(x = x, y = y, fill = IFL_raster), color = "green") +
  # geom_sf(data = amazon_main, color = "blue2", size = 1, alpha = 0.6) +
  # geom_sf(data = indigenous_areas_brazil, fill = "orange", alpha = 0.9, color = "grey10") +
  #### Fires proportional circles (color not mapped)
   # geom_sf(data = points_fire_proportion, aes(size = n, color = pct), 
        ## Smaller dataset (faster plotting)
         # geom_sf(data = fire_sample, color = "red2", size = 0.95,
         # alpha = 0.7, inherit.aes = FALSE) +
#### Amazon biome outline
    # geom_sf(data = amazon_biome, fill = NA, alpha = 0.1) +

  #### South america country borders    
  # geom_sf(data = samer_countries, fill = NA,
    # color = "grey45") +
  #### Fires proprtional circles (colors mapped)
  geom_sf(data = points_fire_proportion, aes(size = n, color = pct_rcl),
          # color = "red2",
          inherit.aes = FALSE) +
  scale_color_manual(
          # in manual scales, one has to define colors, well, manually
          # I can directly access them using viridis' magma-function
          values = colorscale,
          breaks = rev(brks_scale)) +
 
      geom_sf(data = samer_countries, fill = NA,
    # # linetype = "dotted",
    color = "grey40") +
  geom_sf(data = amazon_biome, fill = NA, alpha = 0.1, color = "grey10",
          size = 1) +
  coord_sf(datum = NA, xlim = c(-78, -42), ylim = c(-19, 10)) +
    theme_map() +
  theme(
    legend.position = "none"
  )

plot_firedens

ggsave("Output/Figures/map_firefreq_indigenous.png", plot_firedens,
        dpi = 400, width = 16, height = 12)


```


